Curiosity-driven Exploration of Skill Hierarchies

### 2015-11-06 paper1/paper.tex
xp1 xp2

@@

	The same idea holds for language expressiveness which is based on syntactic hierarchical combinations 
	on a vocabulary, that open infinite semantic possibilities. 
	Greenfield has also argued that this parallel between manipulation and language compositionality can be found in the human ontegenic development
	with combinatorial steps for manipulation and syntax acquired approximately at the same period and in the same order \cite{green}. 
	Also, the author explains that the development of the neural substrates for language and tool use could be an ontogenic homology as first of all
	the same neural computations for hierarchical combinations and their semantics should take place for both modalities, and furthermore 
	experiments with Broca's and Wernicke's aphasics show that hierarchical organization for language and manipulation is linked. 
	Broca's aphasics, who have less syntactic organization of speech were shown to also have problems of representation of the hierarchical organization
	of constructions with blocks, whereas Wernicke's aphasics, whose syntax is normal but speech semantics is impaired, succeed in representing such objects 
	hierarchies.

	Functional MRI experiments by Higuchi et al. have shown that the human's neural substrates for tool use and language is indeed shared in 
	the dorsal BA44 Broca's area \cite{higuchi}, which gives evidence for the similar neural computations used. 
	They furthermore argue that these results supports the hypothesis that tool use have appeared first in primate evolution in F5 area,
	and then the language has developed in humans reusing part of tool use and manipulation neural substrates in human's Broca area, homolog of primate's F5.

@@

	\subsection{Goal of the study}

		\begin{itemize}
			\item Exploring in a structured hierarchy is more efficient than directly from $M$ to $S$.
			\item Which task should I explore now ?
			\item How to choose between different means to explore a given space ? 
			\item How can high-level tasks guide the exploration of lower-level ones ?
			\item How can the system cope with perturbations on some of the forward models ?		
		\end{itemize}

@@

	\subsection{Related work}
		\cite{ugur2014, ugur2015}.

@@

		The model from Fabisch et al. \cite{fabisch2014active} learns in a setting with a discrete task space (called contexts).
		It uses an intrinsic motivation for learning progress, and a Multi-Armed Bandit algorithm (D-UCB) to choose on which context the agent should train for.
		The Upper Confidence Bound algorithm chooses between contexts given their estimated learning progress and the uncertainty of these estimations
		by picking the context with the maximum upper confidence bound.
		In other words, it maximizes the expected reward plus something related to the uncertainty associated with it, selecting either contexts with 
		certain high rewards or ones with uncertain poor reward.
		This algorithm embeds directly a solution the exploration-exploitation trade-off problem as it represents the exploitation 
		of knowledge by the expected progress and the exploration of other solutions by the uncertainty bonus. 
		This algorithm supposes a stationary learning progress on each context so the authors use an adaptation 
		(D-UCB, \cite{kocsis2006discounted}) to encompass non-stationary learning progress.
		% Sliding Window UCB: \cite{garivier2011upper}

		In a fully continuous setting, Mugan et al. \cite{mugan} have developed an algorithm that first learns a qualitative representation of environment states
		and actions in order to then learn the structure of Dynamic Bayesian Networks representing the temporal contingencies of those states and actions.
		In order to choose which action to practice, the authors use the IAC \cite{oudeyer_intrinsic_2007} where the agent 
		is intrinsically motivated to choose actions that are estimated to yield high prediction error progress.

		Here, we will rely more specifically on the SAGG-RIAC architecture \cite{baranes_active_2013}.
		This architecture learns a single mapping between continuous motor and sensori (or task) 
		spaces with a competence-based intrinsic motivation. 
		In our hierarchy of sensorimotor models, each model will be explored using the SAGG-RIAC procedure, but it could be replaced by another one 
		without changing the mechanisms to learn the hierarchy that will be assessed.

@@

\section{Environment}

@@

@@

		We simulate a simple vocal signal controled by pitch and intensity between $-1$ and $1$.
		The pair will engage in helping the robot only if pitch and intensity were sufficiently high ($>0.$).
	%

@@

		The trajectory of the hand and of the end point of the sticks is the sequence of x and y positions at different time points: steps $12$, $25$, $37$ during the movement of $50$ steps.
		The trajectory of the gripper is a sequence of $1$ or $-1$ depending whether the gripper is open or not.
		The pair understanding of the vocal signal is represented as $1$ if intensity and pitch were correct, $-1$ otherwise.
		For each box, the robot receives a $1$ if the object was in the box, and $0$ otherwise, plus the distance between the end position of the object and the center of the box.

		The sensory information thus contains $6$ dimensions for the trajectory of the hand, $1$ for the pair help, $6$ for the trajectory of the end of each stick, $2$ for the end position of the object, and $2$ for each box.
		The total sensory space has $44$ dimensions.

@@

\section{Random Goal Babbling}

@@

		\begin{itemize}
			\item Idea: to compare our simplest algorithm (explore the module with higher progress) in a realistic setting (Hierarchy (a) of Fig. \ref{Hierarchies}) to the control condition where a sensorimotor model is learned
					directly from the whole motor space $M$ to the whole sensori space $S$.

			\item Conditions: Hierarchy (a) vs $M \rightarrow S$, Motor Babbling vs SAGG-Random

			\item Features: MAB on all modules, NN, No TDD.

@@

		\begin{figure*}[!t]
			\centering
			\subfloat[Competences]{\includegraphics[width=6cm]{./include/H0-GB-log6-competences.png}}
			\subfloat[Interests]{\includegraphics[width=6cm]{./include/H0-GB-log6-interests.png}}
			\subfloat[Events]{\includegraphics[width=6cm]{./include/H0-GB-log6-events.png}}
			\caption{Competences, interests and events using H0 with Random Goal Babbling}
			\label{resH0GB}
		\end{figure*}

		\begin{figure*}[!t]
			\centering
			\subfloat[Competences]{\includegraphics[width=6cm]{./include/H1-GB-log5-competences.png}}
			\subfloat[Interests]{\includegraphics[width=6cm]{./include/H1-GB-log5-interests.png}}
			\subfloat[Events]{\includegraphics[width=6cm]{./include/H1-GB-log5-events.png}}
			\caption{Competences, interests and events using H1 with Random Goal Babbling}
			\label{resH1GB}
		\end{figure*}

		\begin{figure*}[!t]
			\centering
			\subfloat[Tool]{\includegraphics[width=6cm]{./include/xp1-n_tool_1.png}}
			\subfloat[Object]{\includegraphics[width=6cm]{./include/xp1-n_obj.png}}
			\subfloat[Box 1]{\includegraphics[width=6cm]{./include/xp1-n_box_1.png}}\\
			\subfloat[Box 2]{\includegraphics[width=6cm]{./include/xp1-n_box_2.png}}
			\subfloat[Box 3]{\includegraphics[width=6cm]{./include/xp1-n_box_3.png}}
			\caption{Number of touch of tool, object and boxes for each 100 iterations' bin.}
			\label{res_xp1_events}
		\end{figure*}

@@

		\begin{itemize}
			\item Idea: to compare the different possibilities to choose the module to explore in the hierarchy: maximizing the progress, maximizing with a bias towards lower-level modules, or use ZPDES,
					with the same hierarchy (a) of Fig. \ref{Hierarchies}.

			\item Conditions: Random module, MAB on all modules, MAB with bias, ZPDES.

			\item Features: Hierarchy (a), SAGG-Random, NN, No TDD.

@@

>>>>

	Curiosity-driven exploration and developmental trajectories...

@@

@@

	Here, we will rely on the SAGG-Random architecture \cite{baranes2010intrinsically}.
	This architecture learns a single mapping between a continuous motor space and and continuous sensory (or task) 
	space by randomly selecting goals in the sensory space. 
	In our hierarchy of sensorimotor models, each model will be explored using the SAGG-Random procedure, but it could be replaced by any other exploration architecture.

	We want to study and compare our exploration strategies in a high-dimensional continuous environment where primitive actions can be reused to influence different perceptual spaces.
	We will use a simulated 2D robotic arm that can interact with tools and objects. 
	The robot will have the possibility to use different tools and to ask for the help of a pair to reach certain objects.	
	Section \ref{sec:env} shows the details of that environment.

	Section \ref{sec:saggrandom} explains in detail the SAGG-Random exploration architecture and shows how it behaves on the designed environment.

	\subsection{Questions of the study}

		\begin{itemize}
			\item Learning a structured hierarchy is more efficient than directly from $M$ to $S$.?
			\item Which task should be explored now ?
			\item How to choose between different means to explore a given space ? 
			\item How can high-level tasks guide the exploration of lower-level ones ?
			\item How can the system cope with perturbations of some of the forward models ?		
		\end{itemize}

@@

	\subsection{Related work}
		Explain \cite{ugur2014, ugur2015, schmerlinggoal}.

@@

		The model from Fabisch et al. \cite{fabisch2014active} learns a discrete task space (tasks are called contexts).
		It uses an intrinsic motivation for learning progress, and a Multi-Armed Bandit algorithm (D-UCB) to choose on which context the agent should train for.
		The Upper Confidence Bound algorithm chooses between contexts given their estimated learning progress and the uncertainty of these estimations
		by picking the context with the maximum upper confidence bound.
		%In other words, it maximizes the expected reward plus something related to the uncertainty associated with it, selecting either contexts with 
		%certain high rewards or ones with uncertain poor reward.
		%This algorithm embeds directly a solution the exploration-exploitation trade-off problem as it represents the exploitation 
		%of knowledge by the expected progress and the exploration of other solutions by the uncertainty bonus. 
		This algorithm supposes a stationary learning progress on each context so the authors use an adaptation 
		\cite{kocsis2006discounted} to encompass non-stationary learning progress.
		% Sliding Window UCB: \cite{garivier2011upper}

		In a fully continuous setting, Mugan et al. \cite{mugan} have developed an algorithm that first learns a qualitative representation of environment states
		and actions in order to then learn the structure of Dynamic Bayesian Networks representing the temporal contingencies of those states and actions.
		In order to choose which action to practice, the authors use the Intelligent-Adaptive-Curiosity algorithm \cite{oudeyer_intrinsic_2007} where the agent 
		is intrinsically motivated to choose actions that are estimated to yield high prediction error progress.

@@

\section{Environment}
\label{sec:env}

@@

	A simpler version of that environment will also be considered, where only the small stick is provided to the robot (environment E1) instead of both (environment E2).

@@

		We simulate a simple vocal signal controled by pitch and intensity between $-1$ and $1$.
		The pair will engage in helping the robot only if pitch and intensity are sufficiently high ($>0$).
	%

@@

		The trajectory of the hand and of the end point of the sticks is the sequence of x and y positions at different time points: steps $12$, $25$, $37$ during the movement of $50$ steps.
		The trajectory of the gripper is a sequence of $1$ or $-1$ depending whether the gripper is open or not.
		The pair understanding of the vocal signal is represented as $1$ if intensity and pitch were correct, $-1$ otherwise.
		The sensory information about the boxes is composed of 2 values. 
		The first one tells which box has the object inside ($-1$ if no one, $-0.25$ if box 1, $0.25$ if box 2, and $0.75$ if box 3).
		The second value is the minimal distance between a box and the object at the end of the movement and after the pair helped.
		%For each box, the robot receives a $1$ if the object was in the box, and $0$ otherwise, plus the distance between the end position of the object and the center of the box.

		The sensory information thus contains $6$ dimensions for the trajectory of the hand, $1$ for the pair help, $6$ for the trajectory of the end of each stick, $2$ for the end position of the object, and $2$ for the boxes.
		The total sensory space has $34$ dimensions if only the small stick is provided to the robot (environment E1), and $40$ dimensions if both are (environment E2).
		%The sensory information thus contains $6$ dimensions for the trajectory of the hand, $1$ for the pair help, $6$ for the trajectory of the end of each stick, $2$ for the end position of the object, and $2$ for each box.
		%The total sensory space has $44$ dimensions.

@@

\section{Random Goal Babbling}
\label{sec:saggrandom}

	Explain SAGG-Random, NN, show results on environment for GB vs MB.
	Explain exploration in SAGG-Random, show competence and progress with different measures (and why its important to measure that).
	The possible perturbation of the standard measure by exploration noise, and a possible fix with the following:

	We have $$||s_g - s|| = ||s_g - s_{NN} + s_{NN} - s_p + s_p - s||$$ where $s_{NN}$ is the nearest neighbor of $s_g$ in the sensorimotor mapping, and is also the sensory prediction of $m$, and $s_p$ is the sensory prediction of $m + \epsilon$.

	When there is no exploration, $\epsilon=0$, so $s_{NN} = s_p$ and $$||s_g - s|| = ||s_g - s_{NN} + s_p - s||$$
	The expression $s_g - s_{NN}$ corresponds to the novelty with respect to the already reached points, and $s_p - s$ to the forward model's prediction error.

	We could also use $||s_g - s_{NN} + s_p - s||$ when $\epsilon > 0$ as the basis of the computation of competence and interest.

@@

		Here we want to compare the learning of a structured hierarchy of SAGG-Random modules with the control condition where one sensorimotor model learns
		directly from the whole motor space $M$ to the whole sensori space $S$. 
		For the hierarchical conditions, we randomly choose the module of the hierarchy that will explore at each iteration.
		We also compare Motor Babbling and Goal Babbling in the hierarchical and in the flat hierarchies.
		We use hierarchy H1 (Fig. \ref{H1}), and environmnent E1.
		Exploration: only the babbling module will add exploration noise.

		We measure the number of times the agent manages to grasp the tool, push the object, and reach a box.
		TODO: exploration of all spaces (hand, tools, object, boxes), competence to reach random goals.

@@

		\begin{figure*}[!t]
			\centering
			\subfloat[Competences]{\includegraphics[width=6cm]{./include/H0-GB-log1-competences.png}}
			\subfloat[Interests]{\includegraphics[width=6cm]{./include/H0-GB-log1-interests.png}}
			\subfloat[Events]{\includegraphics[width=6cm]{./include/H0-GB-log1-events.png}}
			\caption{Competences, interests and events using H0 with Random Goal Babbling}
			\label{resH0GB}
		\end{figure*}

		\begin{figure*}[!t]
			\centering
			\subfloat[Competences]{\includegraphics[width=6cm]{./include/H1-GB-log10-competences.png}}
			\subfloat[Interests]{\includegraphics[width=6cm]{./include/H1-GB-log10-interests.png}}
			\subfloat[Events]{\includegraphics[width=6cm]{./include/H1-GB-log10-events.png}}
			\caption{Competences, interests and events using H1 with Random Goal Babbling}
			\label{resH1GB}
		\end{figure*}

		\begin{figure*}[!t]
			\centering
			\subfloat[Tool]{\includegraphics[width=6cm]{./include/xp1-tool_1.png}}
			\subfloat[Object]{\includegraphics[width=6cm]{./include/xp1-obj.png}}
			\subfloat[Box 1]{\includegraphics[width=6cm]{./include/xp1-box_1.png}}\\
			\subfloat[Box 2]{\includegraphics[width=6cm]{./include/xp1-box_2.png}}
			\subfloat[Box 3]{\includegraphics[width=6cm]{./include/xp1-box_3.png}}
			\caption{Number of touch of tool, object and boxes for each 100 iterations' bin.}
			\label{res_xp1_events}
		\end{figure*}

@@

		Here we want to compare the different possibilities to choose the module to explore in the hierarchy: choosing the one with the max progress (proba proportional to interest), 
		maximizing with a bias towards lower-level modules, or use ZPDES,
		with the same hierarchy H1 (Fig. \ref{H1}), and environmnent E1.

		Conditions: Random module (RD), choose with proba proportional to interest (I), the same with bias toward lower levels (IB), or ZPDES.
		Also try to make all modules add exploration noise (EXPLOA) or only the babbling module (EXPLOB).

		We measure the number of times the agent manages to grasp the tool, push the object, and reach a box.
		TODO: exploration of all spaces (hand, tools, object, boxes), competence to reach random goals.
	%

@@

		See Fig. \ref{resIEXPLOB} and \ref{res_xp2_events}.

		\begin{figure*}[!t]
			\centering
			\subfloat[Competences]{\includegraphics[width=6cm]{./include/xp2-I-EXPLOB-log4-competences.png}}
			\subfloat[Interests]{\includegraphics[width=6cm]{./include/xp2-I-EXPLOB-log4-interests.png}}
			\subfloat[Events]{\includegraphics[width=6cm]{./include/xp2-I-EXPLOB-log4-events.png}}
			\caption{Competences, interests and events using H1 with I-EXPLOB}
			\label{resIEXPLOB}
		\end{figure*}

		\begin{figure*}[!t]
			\centering
			\subfloat[Tool]{\includegraphics[width=6cm]{./include/xp2-tool_1.png}}
			\subfloat[Object]{\includegraphics[width=6cm]{./include/xp2-obj.png}}
			\subfloat[Box 1]{\includegraphics[width=6cm]{./include/xp2-box_1.png}}\\
			\subfloat[Box 2]{\includegraphics[width=6cm]{./include/xp2-box_2.png}}
			\subfloat[Box 3]{\includegraphics[width=6cm]{./include/xp2-box_3.png}}
			\caption{Number of touch of tool, object and boxes for each 100 iterations' bin.}
			\label{res_xp2_events}
		\end{figure*}
